Manuscript title here

### 2024-09-16 manuscript.Rmd
Uppdaterad intro, med tillagd OFI del från projektplan samt även ändrat strukturen från innan och ändrat enligt kommentarerna

Trauma is a major global public health concern in form of death and disability. It affects over four milion peole each year, among those younger populations are the most affected <!-- You can be explicit and say that trauma causes over four million deaths each year. Trauma affects hundreds of millions -->[@noauthor_injuries_nodate]. The initial management of trauma patients once they reach hospital is very time sensitive and error prone [@mackersie_pitfalls_2010]. This combined with the fact that trauma patients in shock have a higher mortality rate, than those not[@vang_shock_2022], makes it important to study and improve the care for trauma patients in shock. <!-- You can write that shock is a major cause of mortality in trauma patients. -->

Shock are generally defined as circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are four types of causes, which are hypovolemic, obstructive, cardiogenic and distributive. These causes can occur alone or in combination. The most usual one during trauma are hypovolemic in form of a large bleeding[@teixeira_preventable_2007]. <!-- Move the information about the types of shock here. -->

Clinically to be able to differentiate the degree of shock, the ATLS classification for major haemorrhage can be used. The classification are divided into four classes, based on the patients: estimated bloodloss, heart frequency (HF), bloodpressure (BP), pulse pressure, respiratory rate (RR), mental status and urine output. However this only suits hypovolemic shocks and not the other types. There are two other classifications, namely RETTS and NEWS2<!-- You can discuss these in your KI paper but you don't need it here in your manuscript-->. These are used in emergency triaging and can be used on other types of shocks, however they are not shock specific classifications. If we look into all three classifications we will be able to find similarities in that all use HF, BP, RR and mental status. (sista två meningarna borde tillhöra metod...)

<!-- Say something about opportunities for improvement here, and why it's important to study these among patients in shock -->

Infoga OFI del och koppla samman

>>>>

Trauma is a major global public health concern. It causes over four milion death and affects hundreds of millions each year, among those younger populations are the most affected [@noauthor_injuries_nodate]. Shock is a major cause of mortality in trauma patients. The definition for shock is circulatory failure leading to insuficient perfusion and oxygenation of tissue and organ[@leech_shock_2023]. There are four types of causes, which are hypovolemic, obstructive, cardiogenic and distributive. These causes can occur alone or in combination. The most usual one during trauma are hypovolemic in form of a large bleeding[@teixeira_preventable_2007].

Clinically to be able to differentiate the degree of shock in context of trauma, the ATLS classification for major haemorrhage can be used. The classification is divided into four classes, based on the patients: estimated bloodloss, heart frequency (HF), bloodpressure (BP), pulse pressure, respiratory rate (RR), mental status and urine output. However, this only suits hypovolemic shocks and not the other types. (Källa) There is other classification for triaging or tracking the clinical deterioration of patients, which can be used for other causes of shocks. The common denominator for these classifications is they are all based on vital parameters.

The initial management of trauma patients once they reach hospital is very time sensitive and error prone [@mackersie_pitfalls_2010]. To improve trauma care quality, healthcare systems identify Opportunities for Improvement (OFI) through structured reviews of patient cases. Multidisciplinary morbidity and mortality reviews play a central role in this process, where clinicians from different specialities together analyse trauma cases to evaluate both the structure and clinical processes involved in trauma care following the Donabedian’s quality of care framework[@albaaj_patient_2023]. For shock patients, timely interventions such as fluid resuscitation/blood transfusion, intervention, OR and radiology are critical, and delays or missteps can result in adverse outcomes (källa). By systematically addressing OFIs, healthcare providers can improve the effectiveness of trauma care, directly impacting mortality and morbidity(källa).
